My three plots:

1.  depth profiles of concentration of nitrogen and carbon vs depth (dual axis plot)

Data wrangling and preparation:

I chose to visualize data observations whose depths were greater than or less than 80 meters below sea level.

The variables of interest in this visualization are :

-   `botid_HOT` : unique water sensor ID

-   `doc_bottle_hot` : Dissolved Organic Carbon (µmol/kg)

-   `don_bottle_hot` : Dissolved Organic Nitrogen (µmol/kg)

-   `ph_bottle_hot` : Bottle pH (1 is most acidic and 14 is most basic)

-   `oxygen_ctd_bottle_hot` : Bottle CTD Oxygen (µmol/kg)

-   `temperature_ctd_bottle_hot` : Bottle CTD Temperature (ITS- 90, a temperature scale based on the Celsius scale)

```{r include = FALSE, message = FALSE, warning = FALSE}

# loading libraries and fonts for plots
library(tidyverse)
library(showtext)
library(here)
library(patchwork)
library(sysfonts)
library(ggtern)
library(lubridate)
library(rcartocolor)
library(scales)
library(ggridges)
font_add_google(name = 'Josefin Sans', family = 'josefin')
font_add_google(name = 'Inter', family = 'inter')

showtext_auto()

# loading in data
nutrient_data <- read_csv(here::here('data','HOT_Bottle_ALOHA.csv'))

# preliminary data cleaning
nutrient_clean <- nutrient_data %>% 
  select(botid_HOT,time,doc_bottle_hot,don_bottle_hot,depth,chl_bottle_hot,ph_bottle_hot, oxygen_ctd_bottle_hot, NO2_NO3_bottle_hot,temperature_ctd_bottle_hot) %>% # selecting columns of interest
  na.omit() %>% # removing NAs
  # only keeping observations with depth less than or equal to 80 m below sea level
  rename(bottle_id = botid_HOT,
         dissolved_c = doc_bottle_hot,
         dissolved_n = don_bottle_hot,
         ph = ph_bottle_hot,
         chlorophyll = chl_bottle_hot,
         oxygen = oxygen_ctd_bottle_hot,
         temperature = temperature_ctd_bottle_hot)

# scaling coefficient
coeff <- 10
# 999933 - Nitrogen 
# 117733 - carbon

# subsetting dataset to only have 80m below sea level; categorizing depths in groups for box plot using case_when
nutrient_clean_2 <- nutrient_clean %>% 
  filter(depth <= 80) %>% 
  mutate(avg_n = mean(dissolved_n),
         avg_c = mean(dissolved_c),
         depth_group = case_when(depth <= 20 ~ 'shallow',
                                 depth <= 40 ~ 'medium',
                                 depth <= 60 ~ 'medium deep',
                                 depth <= 80 ~ 'deep'))
# saving the average value of dissolved nitrogen as a variable
avg_dissolved_n <- nutrient_clean_2$avg_n[1]

# saving the average value of dissolved carbon as a variable 
avg_dissolved_c <- nutrient_clean_2$avg_c[1]
```

My first plot is a visual that can be used in a report or technical documentation. I chose to plot using dual axes in order for the reader to capture the different scalings of the variables used in the graph. I chose to incorporate the horizontal lines in order to help center the reader's attention on the stagnant average concentration of both values. I included the use of color blind friendly colors, along with

```{r message = FALSE}
#..............................first plot........................................


ggplot(nutrient_clean_2, aes(x = depth)) +
  # first line plot is of dissolved nitrogen
  geom_line(aes(y = dissolved_n, color = "Nitrogen")) +
  # adding a horizontal dotted line representing the average dissolved nitrogen       value
  geom_hline(yintercept = avg_dissolved_n,
             color = "lightgray",
             linetype = 'dotted') +
  # adding a horizontal dotted line representing the average dissolved carbon         value
  geom_hline(
    yintercept = avg_dissolved_c / coeff,
    color = 'lightgray',
    linetype = 'dotted'
  ) +
  # second line plot is of dissolved carbon divided by the coefficient in order       for it to show on the same plane as dissolved nitrogen
  geom_line(aes(y = dissolved_c / coeff, color = "Carbon")) +
  # specifying the two axes and labeling
  scale_y_continuous(name = 'Dissolved Nitrogen (µmol/kg)',
                     sec.axis = sec_axis( ~ . * coeff,
                                          name = 'Dissolved Carbon (µmol/kg)')) +
  # using color blind friendly colors to distinguish the different lines
  scale_color_manual(values = c('Nitrogen' = '#999933',
                                'Carbon' = '#117733')) +
  # adding x axis label, title, and citation
  labs(x = 'Depth (m)',
       title = 'Depth Profile Comparison of Dissolved Carbon and Nitrogen',
       tag = "Data Source: Hawaiian Ocean Time-series (HOT)") +
  # reversing the x axis values so that the top is 0 and the bottom is 80
  scale_x_reverse() +
  # adjusting the x and y axis title fonts and sizes
  # removing the legend title, panel border, panel grid major and minor lines
  # specifying the tag position for the citation, specifying legend text
  theme(
    aspect.ratio = 2,
    axis.title.x = element_text(family = 'inter', size = 14),
    axis.title.y = element_text(family = 'inter', size = 14),
    plot.title = element_text(
      family = 'josefin',
      size = 18,
      hjust = .5
    ),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    legend.title = element_blank(),
    legend.text = element_text(family = 'inter', size = 14),
    plot.tag.position = c(0.9999, 0),
    legend.key = element_blank()
  ) +
  # adding annotations to specify average concentrations for the horizontal lines     generated 
  annotate(
    "text",
    x = 60,
    y = 4.5,
    label = 'Average concentration: 5.8 (µmol/kg)',
    angle = -90,
    size = 4
  ) +
  annotate(
    "text",
    x = 60,
    y = 11.5,
    label = 'Average concentration: 95.1 (µmol/kg)',
    angle = -90,
    size = 4
  ) +
  # flipping axes
  coord_flip()

```

2.  Depth profile of chlorophyll and its relationship to pH and dissolved oxygen

```{r warning = FALSE, message = FALSE}
# scaling pH and dissolved oxygen to match chlorophyll
o_coeff <- 100
ph_coeff <- 10


cols <- c('#edf8e9', '#bae4b3', "#74c476", '#238b45')
ggplot(data = nutrient_clean_2, aes(y = oxygen, x = depth)) +
  geom_boxplot(
    aes(fill = depth_group),
    alpha = 0.75,
    outlier.color =  "#54278f",
    outlier.shape = 17,
    outlier.size = 2
  ) +
  geom_point(aes(x = depth, y = oxygen), alpha = 0.25) +
  scale_fill_manual(
    values = rev(cols),
    labels = c(
      '(60 < depth <80) ',
      '(40 < depth < 60)',
      '(20 < depth < 40)',
      ' (20 < depth < 0)'
    )
  ) +
  scale_x_reverse() +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 8)) +
  labs(
    fill = 'Water Depth (m)',
    y = 'Bottle CTD Oxygen (µmol/kg)',
    x = 'Depth (m)',
    title = 'Oxygen Concentration and its Relationship \nto Depth and pH within the Epipelagic Zone',
    tag = "Data Source: Hawaiian Ocean Time-series (HOT)"
  ) +
  coord_flip() +
  theme(
    axis.title.x = element_text(
      family = 'inter',
      size = 14,
      margin = margin(10, 0, 0, 0)
    ),
    axis.title.y = element_text(
      family = 'inter',
      size = 14,
      margin = margin(0, 10, 10, 0)
    ),
    plot.title = element_text(
      family = 'josefin',
      size = 18,
      hjust = .5,
      lineheight = 1,
      margin = margin(0, 0, 10, 0)
    ),
    legend.title = element_text(
      family = 'inter',
      size = 16,
      hjust = .5
    ),
    legend.text = element_text(family = 'inter', size = 12),
    axis.text = element_text(family = 'inter', size = 12),
    legend.key = element_blank(),
    legend.background = element_rect(linewidth = 0.05, color = 'black'),
    panel.background = element_blank(),
    axis.line = element_line(colour = "darkgrey"),
    plot.tag.position = 'bottom',
    plot.tag = element_text(
      family = 'inter',
      size = 10,
      face = 'italic',
      margin = margin(10, 0, 0, 0)
    )
  ) +
  annotate(
    'text',
    x = 24,
    y = 166,
    label = 'Outliers are represented with a purple triangle.',
    hjust = 0
  ) +
  annotate(
    'text',
    x = 10,
    y = 166,
    label = 'Oxygen concentration was measured using CTD sensors,\n which stands for conductivity, temperature, and density.',
    hjust = 0
  ) +
  theme(text = element_text(margin = margin(0.5,0,0,0)))

```

3.  line graph with element_rect()

```{r}
arrow_df <- data.frame(x1 = 45,
                       y1 = 8,
                       x2 = 56,
                       y2 = 8.76)
point <- data.frame(y = 9.36, x = 57.9)
ggplot(data = nutrient_clean_2) +
  geom_point(
    aes(x = depth, y = dissolved_n, size = temperature),
    color = '#09555c',
    alpha = 0.5
  ) +
  geom_point(
    data = point,
    aes(x = x, y = y),
    color = 'red',
    size = 4,
    alpha = 0.45
  ) +
  labs(x = 'Depth (meters)',
       y = ' Concentration (µmol/kg)',
       title = 'Depth Profile of Dissolved Nitrogen',
       size = 'Temperature (°C)') +
  scale_x_reverse() +
  scale_y_continuous(breaks = seq(4, 10, 1)) +
  coord_flip() +
  geom_smooth(
    aes(x = depth, y = dissolved_n),
    se = FALSE,
    color = 'black',
    linewidth = 0.45
  ) +
  theme_bw() +
  theme(
    axis.title.x = element_text(family = 'inter', size = 14),
    axis.title.y = element_text(family = 'inter', size = 14),
    plot.title = element_text(
      family = 'josefin',
      size = 18,
      hjust = .5
    ),
    axis.text.x = element_text(family = 'inter', size = 13),
    axis.text.y = element_text(family = 'inter', size = 13),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = '#E5E4E2', size = 0.25),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = '#E5E4E2', size = 0.25),
    plot.background = element_rect(fill = "#c9e9f6"),
    legend.text = element_text(family = 'inter', size = 12),
    legend.background = element_rect(fill = "#c9e9f6"),
    legend.key = element_blank()
  ) +
  geom_curve(aes(
    x = x1,
    y = y1,
    xend = x2,
    yend = y2
  ),
  data = arrow_df,
  arrow = arrow(length = unit(0.03, 'npc'))) +
  annotate(
    "rect",
    xmin = 50,
    xmax = 70,
    ymin = 8.75,
    ymax = 10,
    size = .05,
    fill = 'lightgrey' ,
    alpha = 0.25
  ) +
  annotate(
    'text',
    x = 36,
    y = 9.05,
    label = "Descending from the surface, the concentration of Nitrogen\n spikes to 9.36 µmol/kg. At ~57.9 meters below surface\n level, the pH from this oceanographic sample was 8.06.",
    fontface = 'italic',
    hjust = 0.5
  )
```
